When does vapor pressure deficit drive or reduce evapotranspiration?

### 2018-03-29 doc/paper/vpd_et_paper.tex
refactored and back to writing

\subsection{Functional form of ET dependence on VPD and its relation to uWUE}

Th theory described in Section \ref{sign_term} indicates that for a given $uWUE$ and $g_1$, the ET dependence on VPD should be concave upward. In other words, there should be some local minima in ET at a critical VPD$_{crit}$, assuming the scaling term (e.g. aerodynamic conductance and $\Delta$) is held fixed. This result warrants futher investigation, because to our knowledge no one has derived the theoretical ecosystem-scale relationship between ET and VPD while controlling for enivironmental conditiions. In particular, from personal communication, there is an apparent lack of consensus over whether the shape of the ET-VPD curve should be concave upward (our result) or concave downward. Given that understanding the ET-VPD relationship of the one-way plant response is absolutely fundamental for hypothesizing about any feedbacks between the land surface and the atmopshere, we analyze why our derived ET-VPD relationship is concave upward, particularly with respect to our use of $uWUE$ to make explicit the relationship betwen GPP, ET, and VPD.

>>>>

\subsection{Functional form of ET dependence on VPD and its relation to VPD dependence}

Th theory described in Section \ref{sign_term} indicates that for a given $uWUE$ and $g_1$, the ET dependence on VPD should be concave upward. In other words, there should be some local minima in ET at a critical VPD$_{crit}$, assuming the scaling term (e.g. aerodynamic conductance and $\Delta$) is held fixed. This result warrants futher investigation, because to our knowledge no one has derived the theoretical ecosystem-scale relationship between ET and VPD while controlling for enivironmental conditiions. In particular, from personal communication, there is an apparent lack of consensus over whether the shape of the ET-VPD curve should be concave upward (our result) or concave downward. Given that understanding the ET-VPD relationship of the one-way plant response is fundamental to hypothesizing about any feedbacks between the land surface and the atmopshere, we analyze why our derived ET-VPD relationship is concave upward, particularly with respect to the exponent of VPD depedence in $uWUE$ and the Medlyn unified stomatal conductace model.

There is a theoretical basis for the square root VPD dependence in both the stomatal conducatance model and $uWUE$ based on the assumption that stomata behave to maximize carbon gain while minimizing water loss, which observations also generally support \citep{Lloyd_1991, MEDLYN_2011, Lin_2015, Zhou_2014, Zhou_2015, Medlyn_2017}. However, purely empiriacle results which fit the exponent of the VPD dependence have shown that it may vary from 1/2, suggesting that stomata, as well as ecosystem-scale quantities based on stomata theroy,  may not always function optimally \citep{Zhou_2015, Lin_2018}. Specifically with regards to $uWUE$, one would not expect that this ecosystem scale WUE quantity will responde to VPD exactly analgalously to stomata. Direct soil evaporation's contributions to ET should shift the exponential dependence of VPD to be greater than 1/2. In other words, direct soil evaporation causes WUE to decrease more strongly with increasing VPD, especially in well-watered conditions. \citet{Zhou_2015}'s results corroborate this: they found a mean imprically fit exponential VPD dependence of 0.55 istead of 0.5 for Ameriflux sites. Given that these recent results on the relatioship between VPD, GPP, and ET form the backbone of our analysis and are what allowed us to derive an explicit expression for ET for the first time (Equation \ref{et}), we will analyze if and how assumptions about the exonent of the VPD dependence impacts the shape of the ET-VPD dependence. By introducing $n$ and $m$ we can free our stomatal conductance model from assumptions about VPD dependence:
\begin{linenomath*}
  \begin{equation}
    g_s = \frac{R \, T}{P} 1.6 \left(1 + \frac{g_1}{VPD^m}\right) \frac{*WUE \; ET}{c_a \; VPD^n},
    \label{m_n}
  \end{equation}
\end{linenomath*}
where:
\[*WUE = \frac{GPP}{ET}VPD^n.\]
To determine how the exponent $n$ and $m$ alter the shape of the ET-VPD dependence we find the roots of the second deriviative of ET, using Equation \ref{m_n} for stomatal conductance, with respect to VPD:
\input{d2_solutions.tex}
